---
title: "xxx"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
date:  "`r Sys.Date()`"
---

### add something new, why not work


## 1. `tidymodels` ecosystem

`tidymodels` provide a complete framework for 

* Data resampling, i.e. training and testing split (`resample``)

* Feature engineering (`recipes``)

* Model fitting (`parsnip`)

* Model tuning (`tune` and `dials`)

* Model evaluation (`yardstick`)

## 2. Basic: modeling in `tidyverse`

### 2.1 `tidyverse` modeling workflow

1. Make a __list column__ `nest()`

2. Work with __list columns__ `map()`

`map(.x, .f = ~mean(.x))`
Can work with `broom`, `Metrics`, `rsample` etc.


3. Simplify the __list_columns__ `unnest()`

4. model presentations `broom::tidy()`

 *  `tidy()` for model cofficients
 
 *  `glance()` one row summary of the model, e.g. R^2
 
 *  `augment()` adds prediction column to the original data
 
```{r, cache=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)

gapminder = readRDS("D:/Dropbox/Coursework/Machine Learning in R/Machine Learning in the Tidyverse/datasets/gapminder.rds")

## Step one is to make a list column based on the grouping variable
nested <- gapminder %>% group_by(country) %>% nest()

## 1. Example 1) calculating the mean of a column 
nested %>% 
  mutate(pop_mean = map(data, ~mean(.x$population))) %>% 
  unnest(pop_mean)

## 2. Example 2) building a model with map()

### 2.1 Coefficients

lm_reg_coef = nested %>%
  mutate(model = map(data, ~lm(formula = life_expectancy~year,data = .x))) %>% 
  mutate(coef = map(model, ~tidy(.x))) %>%
  unnest(coef)

### 2.2 Summary statistics

lm_reg_inference = nested %>%
  mutate(model = map(data, ~lm(formula = life_expectancy~year,data = .x))) %>% 
  mutate(specs = map(model, ~glance(.x))) %>%
  unnest(specs)

### 2.3 augumented dataframes with predicted values, can directly be used for checking prediction

lm_reg_pred =  nested %>%
  mutate(model = map(data, ~lm(formula = life_expectancy~year,data = .x))) %>%
  mutate(augmented = map(model, ~augment(.x))) %>%
  unnest(augmented)

lm_reg_pred %>% filter(country == "Italy") %>%
  ggplot(aes(x = year, y = life_expectancy)) +
  geom_point() +
  geom_line(aes(y = .fitted), color = "red")



```

### 2.2 The sampling workfolow

__Test and train split__

```{r, cache=TRUE, message=FALSE, warning=FALSE}
gap_split = initial_split(gapminder, prop = 0.75)
training_date = training(gap_split)
testing_data = testing(gap_split)
```

__Cross Validation Creation__
```{r, cache=TRUE, message=FALSE, warning=FALSE}
cv_split = vfold_cv(training_data, v = 3)
cv_data = cv_split %>% 
  mutate(train = map(splits, ~training(.x)),
         validate = map(splits, ~testing(.x)))
```